\thetitle

### 2008-01-21 models.tex
Various edits:
  - models.rb:
  	- removed unnecessary #leaves method from Object, Array and Hash
  	- turn "*" at end of line into footnote mark
  - models.yaml:
    - regularized wording of Parametric model entries
    - added footnote markers ("*") at end of fitted Parametric entries
  - paper.tex:
    - changed title
    - work on introduction
    - added footnote text to models table (parametrics fitted with MLE)
  - references.bib:
    - added back Royer01 entry that got lost in the shuffle

\raisebox{1.5pt}{$\centerdot$} overrides all distributions with per-node versions

@@

\raisebox{1.5pt}{$\centerdot$} generic flows assigned to nodes by sampling \\
\raisebox{1.5pt}{$\centerdot$} session start time assigned from pool of start times

@@

\raisebox{1.5pt}{$\centerdot$} generic flows assigned to nodes according to BiPareto model \\
\raisebox{1.5pt}{$\centerdot$} session start time assigned according to Poisson arrival model

@@

\raisebox{1.5pt}{$\centerdot$} flow size \& start time taken from trace \\

@@

\raisebox{1.5pt}{$\centerdot$} flow size is sampled from BiPareto model fitted to inherited distribution \\
\raisebox{1.5pt}{$\centerdot$} start time relative to previous session flow sampled Lognormal fitted to inherited \\

@@

\raisebox{1.5pt}{$\centerdot$} packet sizes and intervals are sampled from inherited distributions

@@

\raisebox{1.5pt}{$\centerdot$} packet sizes and intervals are drawn from fitted Exponential distributions

>>>>

\raisebox{1.5pt}{$\centerdot$} overrides all distributions with node-specific versions

@@

\raisebox{1.5pt}{$\centerdot$} generic flows assigned to nodes by sampling [TODO---more detail] \\
\raisebox{1.5pt}{$\centerdot$} session start time randomly assigned from pool of trace session start times

@@

\raisebox{1.5pt}{$\centerdot$} generic flows assigned to nodes according to fitted BiPareto model\raisebox{1pt}{\footnotesize*} \\
\raisebox{1.5pt}{$\centerdot$} session start time assigned according to Poisson arrival model\raisebox{1pt}{\footnotesize*}

@@

\raisebox{1.5pt}{$\centerdot$} flow size and start time taken from trace \\

@@

\raisebox{1.5pt}{$\centerdot$} flow size is sampled from fitted BiPareto model\raisebox{1pt}{\footnotesize*} \\
\raisebox{1.5pt}{$\centerdot$} start time relative to previous flow in session sampled from fitted Lognormal model\raisebox{1pt}{\footnotesize*} \\

@@

\raisebox{1.5pt}{$\centerdot$} packet sizes and intervals sampled from inherited distributions

@@

\raisebox{1.5pt}{$\centerdot$} packet sizes and intervals drawn from fitted Exponential distributions\raisebox{1pt}{\footnotesize*}

### 2008-01-21 paper.tex
Finally tweaked the models & properties tables to satisfaction.
Minor tweaks.
Various edits:
  - models.rb:
  	- removed unnecessary #leaves method from Object, Array and Hash
  	- turn "*" at end of line into footnote mark
  - models.yaml:
    - regularized wording of Parametric model entries
    - added footnote markers ("*") at end of fitted Parametric entries
  - paper.tex:
    - changed title
    - work on introduction
    - added footnote text to models table (parametrics fitted with MLE)
  - references.bib:
    - added back Royer01 entry that got lost in the shuffle

\newcommand{\thetitle}{Wireless Traffic: The Failure of CBR Modeling}

\usepackage[pdftex]{graphicx}
\usepackage[labelfont=bf,small]{caption}
\usepackage[font=small,labelfont=bf,position=top,nearskip=0em]{subfig}
\usepackage{cite,amsmath,amssymb,rotating,multirow,bigstrut,url,wrapfig}
\usepackage[hyperfigures,bookmarks,bookmarksopen,bookmarksnumbered,colorlinks,linkcolor=black,citecolor=black,filecolor=blue,menucolor=black,pagecolor=blue,frenchlinks=true,pdftitle={\thetitle}]{hyperref}

@@

\title{
\thetitle\\
\vspace{0.2em}
\LARGE{On The Inadequacy of Uniform Modeling for WLAN Workloads}
}
\author{
{\large{Stefan~Karpinski, Elizabeth~M.~Belding, Kevin~C.~Almeroth}}\\
\textit{\{sgk,ebelding,almeroth\}@cs.ucsb.edu}\vspace{0.5em}\\
Department of Computer Science\\
University of California, Santa Barbara
\vspace{-0.5em}
}

@@

\vspace{1em}
\section{Introduction}\label{sec:intro}

@@

\begin{table}[t]
\noindent
\begin{tabular}[t]{|l|}
\hline
\begin{minipage}[l]{3.315in}
\begin{small}
\vspace{0.8em}
\input{properties.tex}
\vspace{0.7em}
\end{small}
\end{minipage}
\\\hline
\end{tabular}
\caption{Summary of properties defined for the hierarchy of object types in the traffic object model ({\footnotesize{TOM}}).}
\end{table}

\begin{table}[t]
\noindent
\begin{tabular}[t]{|l|}
\hline
\begin{minipage}[l]{3.315in}
\begin{small}
\vspace{0.8em}
\input{models.tex}
\vspace{0.7em}
\end{small}
\end{minipage}
\\\hline
\end{tabular}
\caption{Summary of how the models at each behavioral level override or inherit properties in the traffic object model ({\footnotesize{TOM}}).}
\end{table}

@@

The rest of the paper is organized as follows. In Section~\ref{sec:motivation} we present motivation and related work. Our experimental and analytical methodology is presented in Section~\ref{sec:methodology}, while the results of our experiments are explained and analyzed in Section~\ref{sec:results}. The ramifications of these results are discussed in Section~\ref{sec:discussion}. Finally, in Section~\ref{sec:conclusions}, we conclude with a discussion of how this research may be applied to current wireless studies, and how it points the way to better traffic models for the future.

@@

Paxson and Floyd observed in~\cite{Paxson95} that the interplay between end-point behavior and the network conditions is inherently \textit{closed-loop} in the sense that it is potentially affected by complex feedback. Traffic models typically attempt to preserve the closed-loop behavior of network traffic~\cite{Avallone04,Hernandez06}. This presents a fundamental difficulty, however, in that it presumes that we know the intent of end-points: what \textit{would} they have done under different conditions? While we can speculate about what an individual node might hypothetically do, we currently do not understand the impact of the full-network traffic pattern upon performance metrics at all---even without trying to account for hypothetical reactions to alternate situations. Especially in the wireless setting, a fuller understanding of total network behavior must be reached before we can sensibly tackle the complexities of multi-level behavioral feedback. Accordingly, in this paper we attempt to provide a first-order approximation of complete network behavior by studying the response of performance metrics to \textit{open-loop} traffic models without multi-level feedback. It is important to realize that while this does not provide a final picture, we currently lack even a first-order understanding of the effect of different workloads on performance. This first-order understanding is an essential initial step.

@@

There is a large and diverse body of work on traffic analysis, modeling, and generation~\cite{Paxson95,Paxson96,Sommers04,Avallone04,Hernandez06}. We are only able to discuss a small, but hopefully representative sampling of this work. Almost all of the traffic generation work has focused on wide-area Internet backbone traffic.
%Because of the comparatively unlimited capacity of wired networks, traffic workload generally does not offer interesting technical challenges at the edge of the network.
The two most prominent traffic generation frameworks are Harpoon and \caps{D-ITG}. Harpoon~\cite{Sommers04} uses a traffic trace for self-training, and can subsequently generate synthetic traffic with certain statistical properties based on the original trace. The properties reproduced are the empirical distributions of the following: ``file size, inter-connection time, source and destination IP ranges, number of active sessions.'' There is no criterion proposed to determine whether these properties characterize the original traffic adequately---we can only hope that this approximation is good enough. For many purposes, it likely is sufficient; in particular, it is probably appropriate for the intended use in generating traffic for Internet backbone simulations. Wireless networks, however, are particularly sensitive to workload conditions, and sampling from a limited set of empirical distributions does not suffice to reproduce realistic network-wide traffic.

@@

The 24-hour trace is split into 144 10-minute segments, each of which serves as the basis for a set of simulations using different traffic models. The traffic models range from a completely realistic trace-driven model, to a standard \caps{CBR} traffic model. Various partially synthetic intermediate models, described in Section~\ref{sec:models-metrics}, are simulated to study the impact of different aspects of traffic behavior on network performance. To preserve the fairness of the performance comparison, we keep as many features as possible constant across different traffic models. The traffic generated by each synthetic model preserves as many characteristics from the original trace as possible, within the constraints of the model. Moreover, the following features are preserved across all models: the numbers of wireless nodes, the number of flows, the number of application-initiated data units sent, the total bytes of application data sent, and the average flow duration (and therefore the average data rate).

@@

The simulations described in Section~\ref{sec:models-metrics} provide us with the raw data to compare performance metrics for synthetic traffic models with those for real traffic traces. To assess the realism of these models, however, we need a measure of how inaccurate the synthetic performance values are when compared to the real values. Let $x$ be the value of a performance metric using real traffic, and $y$ the value of the same metric using an alternate traffic model, $\M$. Some common measures of error are the difference [$y-x$], the ratio [$y/x$], and the standard error [$(y-x)/x$]. These are all reasonable measures of error; but which is most appropriate for assessing the realism of performance metrics? Instead of picking one arbitrarily, we will first consider the properties that an ideal error function should have, and then use those properties to determine the best measure of error. Moreover, we will show that the unique measure of error that exhibits these ideal properties is the  \textit{log-ratio} of metric values:
\begin{align}
\log(y/x)=\log(y)-\log(x).
\end{align}
In this discussion, $E(x,y)$ is a generic error function applied to the synthetic value, $y$, with respect to the real value, $x$.

@@

Our simulation results are summarized in Figure~\ref{fig:boxplots}. Each subfigure shows a single performance metric. The distribution of errors for each traffic model is visualized with a box-and-whisker plot. These plots allow immediate assessment of realism: a good traffic model should have log-ratio values that are tightly clustered around the center, with a small, evenly balanced box. Additionally, the mean and median markers should be close to the center. Complementing the visual display of summary statistics, Figure~\ref{fig:boxplots} also lists three $p$-values to the right of each box-and-whisker plot. These are, in order, the $p$-values for the median, mean and 3-mean tests described in Section~\ref{sec:statistical-equivalence}. Each test catches a different type of unrealistic statistical behavior.

@@

%\pagebreak
% Hernandez06: ``complete methodology for reproducing the traffic observed on a network link in a closed-loop manner, and proposed a number of metrics for studying the realism of the generated traffic.''
\bibliography{IEEE,references}

>>>>

\newcommand{\thetitle}{LAN Workload: The Failure of Uniform Modeling}

\usepackage[pdftex]{graphicx}
\usepackage[labelfont=bf,small]{caption}
\usepackage[font=small,labelfont=bf,position=top,nearskip=0em]{subfig}
\usepackage{cite,amsmath,amssymb,rotating,multirow,bigstrut,url,wrapfig}
\usepackage[hyperfigures,bookmarks,bookmarksopen,bookmarksnumbered,colorlinks,linkcolor=black,citecolor=black,filecolor=blue,menucolor=black,pagecolor=blue,frenchlinks=true,pdftitle={\thetitle}]{hyperref}
\usepackage{array}

@@

\title{
\thetitle%\\
%\vspace{0.2em}
%\LARGE{On The Inadequacy of Uniform Modeling for WLAN Workloads}
}
\author{
{\large{Stefan~Karpinski, Elizabeth~M.~Belding, Kevin~C.~Almeroth}}\\
\textit{\{sgk,ebelding,almeroth\}@cs.ucsb.edu}\vspace{0.5em}\\
Department of Computer Science\\
University of California, Santa Barbara
\vspace{-0.5em}
}

\newcommand{\Section}[1]{\S\ref{sec:#1}}

@@

\section{Introduction}\label{sec:intro}

More than twenty years of research in analysis and modeling of network traffic have yielded tremendous advances in our understanding of the complex behaviors that emerge from the interaction of millions of humans and billions of computers on the Internet and in local area networks~(\caps{LAN}s). The rate of packet arrivals at routers on the Internet backbone is known to exhibit complex, self-similar patterns with long-range dependence~\cite{Paxson95,Paxson96,Willinger98}. Other research~\cite{Hernandez06:dissertation} has demonstrated how to accurately model the behavior of a single flow of packets between two application endpoints, including ``closed-loop'' feedback (see \Section{related-work}) induced by the ubiquitous transmission control protocol~(\caps{TCP}). Recently, the high-level statistical properties of wireless \caps{LAN} traffic have been rigorously modeled with a variety of parametric distributions~\cite{Hernandez06:wlan-traffic}.

Despite great breakthroughs in understanding and methodology, the networking community remains without a cookbook for generating realistic \caps{LAN} workloads. Worse still, it is unclear what ``realistic'' should mean in this context. Foremost, a clear and testable definition of realism for traffic models is needed. In the absence of such a definition, it remains standard in wireless experiments to generate traffic using the simplest possible model: all nodes, flows and packets are assumed to behave uniformly. Flows are commonly modeled as constant bit-rate (\caps{CBR}) streams of identical packets sent at an invariant rate; the flows in an experiment are commonly given the same duration and data rate; nodes in the network are often selected as flow endpoints randomly or arbitrarily. While more complex approaches have been proposed~\cite{xxx}, the simplest, na\"ive models remain standard because no objective criterion exists by which to compellingly argue that other models are any better. It is simply unclear that simplistic models do any harm, or how, if at all, complex models provide benefit.% in exchange for their increased complexity.

%In this paper, we propose objective criteria by which to judge whether

@@

\begin{table}[t]
\vspace{0.7em}
\begin{tabular}[t]{|l|}
\hline
\begin{minipage}[l]{3.315in}
\begin{small}
\vspace{0.6em}
\input{properties.tex}
\vspace{0.6em}
\end{small}
\end{minipage}
\\\hline
\end{tabular}
\caption{Summary of classes and their attributes in the proposed Traffic Object Model ({\footnotesize{TOM}}) for {\footnotesize{WLAN}} workload generation.}
\vspace{-2em}
\end{table}

\begin{table*}[t]
\begin{center}
\small
\input{models.tex}
\footnotesize\\[0.5em]
\raisebox{1pt}{*}Parametric models fitted to inherited empirical distributions using maximum likelihood estimation ({\scriptsize{MLE}}).
\end{center}
\vspace{-0.75em}
\caption{Summary of how the models at each behavioral level override or inherit properties in the Traffic Object Model ({\footnotesize{TOM}}).}
\vspace{-1.75em}
\end{table*}

@@

The rest of the paper is organized as follows. In \Section{motivation} we present motivation and related work. Our experimental and analytical methodology is presented in \Section{methodology}, while the results of our experiments are explained and analyzed in \Section{results}. The ramifications of these results are discussed in \Section{discussion}. Finally, in \Section{conclusions}, we conclude with a discussion of how this research may be applied to current wireless studies, and how it points the way to better traffic models for the future.

@@

Paxson and Floyd observed in~\cite{Paxson95} that the interplay between end-point behavior and the network conditions is inherently \textit{closed-loop} in the sense that it is potentially affected by complex feedback. Traffic models typically attempt to preserve the closed-loop behavior of network traffic~\cite{Avallone04,Hernandez06:dissertation}. This presents a fundamental difficulty, however, in that it presumes that we know the intent of end-points: what \textit{would} they have done under different conditions? While we can speculate about what an individual node might hypothetically do, we currently do not understand the impact of the full-network traffic pattern upon performance metrics at all---even without trying to account for hypothetical reactions to alternate situations. Especially in the wireless setting, a fuller understanding of total network behavior must be reached before we can sensibly tackle the complexities of multi-level behavioral feedback. Accordingly, in this paper we attempt to provide a first-order approximation of complete network behavior by studying the response of performance metrics to \textit{open-loop} traffic models without multi-level feedback. It is important to realize that while this does not provide a final picture, we currently lack even a first-order understanding of the effect of different workloads on performance. This first-order understanding is an essential initial step.

@@

There is a large and diverse body of work on traffic analysis, modeling, and generation~\cite{Paxson95,Paxson96,Sommers04,Avallone04,Hernandez06:dissertation}. We are only able to discuss a small, but hopefully representative sampling of this work. Almost all of the traffic generation work has focused on wide-area Internet backbone traffic.
%Because of the comparatively unlimited capacity of wired networks, traffic workload generally does not offer interesting technical challenges at the edge of the network.
The two most prominent traffic generation frameworks are Harpoon and \caps{D-ITG}. Harpoon~\cite{Sommers04} uses a traffic trace for self-training, and can subsequently generate synthetic traffic with certain statistical properties based on the original trace. The properties reproduced are the empirical distributions of the following: ``file size, inter-connection time, source and destination IP ranges, number of active sessions.'' There is no criterion proposed to determine whether these properties characterize the original traffic adequately---we can only hope that this approximation is good enough. For many purposes, it likely is sufficient; in particular, it is probably appropriate for the intended use in generating traffic for Internet backbone simulations. Wireless networks, however, are particularly sensitive to workload conditions, and sampling from a limited set of empirical distributions does not suffice to reproduce realistic network-wide traffic.

@@

The 24-hour trace is split into 144 10-minute segments, each of which serves as the basis for a set of simulations using different traffic models. The traffic models range from a completely realistic trace-driven model, to a standard \caps{CBR} traffic model. Various partially synthetic intermediate models, described in \Section{models-metrics}, are simulated to study the impact of different aspects of traffic behavior on network performance. To preserve the fairness of the performance comparison, we keep as many features as possible constant across different traffic models. The traffic generated by each synthetic model preserves as many characteristics from the original trace as possible, within the constraints of the model. Moreover, the following features are preserved across all models: the numbers of wireless nodes, the number of flows, the number of application-initiated data units sent, the total bytes of application data sent, and the average flow duration (and therefore the average data rate).

@@

The simulations described in \Section{models-metrics} provide us with the raw data to compare performance metrics for synthetic traffic models with those for real traffic traces. To assess the realism of these models, however, we need a measure of how inaccurate the synthetic performance values are when compared to the real values. Let $x$ be the value of a performance metric using real traffic, and $y$ the value of the same metric using an alternate traffic model, $\M$. Some common measures of error are the difference [$y-x$], the ratio [$y/x$], and the standard error [$(y-x)/x$]. These are all reasonable measures of error; but which is most appropriate for assessing the realism of performance metrics? Instead of picking one arbitrarily, we will first consider the properties that an ideal error function should have, and then use those properties to determine the best measure of error. Moreover, we will show that the unique measure of error that exhibits these ideal properties is the  \textit{log-ratio} of metric values:
\begin{align}
\log(y/x)=\log(y)-\log(x).
\end{align}
In this discussion, $E(x,y)$ is a generic error function applied to the synthetic value, $y$, with respect to the real value, $x$.

@@

Our simulation results are summarized in Figure~\ref{fig:boxplots}. Each subfigure shows a single performance metric. The distribution of errors for each traffic model is visualized with a box-and-whisker plot. These plots allow immediate assessment of realism: a good traffic model should have log-ratio values that are tightly clustered around the center, with a small, evenly balanced box. Additionally, the mean and median markers should be close to the center. Complementing the visual display of summary statistics, Figure~\ref{fig:boxplots} also lists three $p$-values to the right of each box-and-whisker plot. These are, in order, the $p$-values for the median, mean and 3-mean tests described in \Section{statistical-equivalence}. Each test catches a different type of unrealistic statistical behavior.

@@

%\pagebreak
% Hernandez06:dissertation: ``complete methodology for reproducing the traffic observed on a network link in a closed-loop manner, and proposed a number of metrics for studying the realism of the generated traffic.''
\bibliography{IEEE,references}
